### 2025-05-03 conteudo/methods/methods.tex
add some work from 03/05

The ethnographic study happened in two phases. 
In the first phase, the author of this research explored the process of mitigating
a code duplication found in the AMD Display Driver and sent it as a PATCH to iteract with the
Linux community and get their feedback. On a second phase, we proposed to students on the 
course Free Software Development at University São Paulo to mitigate the code duplications found 
by the tool in the IIO and AMD Display Driver and pass the process of send the mitigations as PATCH.
(DOUBLE CHECK IF THE STUDENTS HAVE SENT THE PATCHES).

I

@@

%
Given the vast number of lines and components in the Linux kernel, analyzing every duplication is impractical. We focused our scope on the AMD Display driver, which, as shown in Figure \ref{fig:relatory_ex}, contains over 20,000 lines of duplicated code at a similarity metric of 100\%.

@@

Ethnography is a recognized qualitative empirical method for understanding people, 
their cultures, and their work practices \citep{bookethno}. It allows researchers 
to gain insights into values, beliefs, and ideas from the perspective of community 
members \citep{ethnosoft}. In our research, refactoring code duplications without 
community interaction would prevent us from understanding whether Linux maintainers 
would accept our approach or identify their concerns. Thus, we will use ethnography, 
positioning ourselves as Linux kernel contributors focused on removing identified 
code duplications.

This section outlines our approach for mitigatin code duplication in the Linux kernel
and describes the ethnographic methodology for engagin with the Linux community. Firstly, 
we present the participant observation research method explored by the author. Then, 
we present the methodolody of the experiment done with the students.

@@

@@

\todo[inline]{Rsearch with students does not look good.}

@@

For the first alternative, one of the teaching assistants used the Arkanjo tool in the IIO subsystem and
listed interesting duplications that the groups can approach, and it is up to the students on how
to refactor the code to mitigate the duplications and pass through the process to send patches to the IIO 
maintainers to review. For the second alternative, we propose a similar experience then the 
participant observation method, where the students pass through the whole process of executing the Arkanjo tool
in the AMD display driver, analyze the duplications found, create a patch to mitigate a duplication and sent
the patch to the AMD display driver's maintainers.

@@

For the teaching assistant and the student that has executed the Arkanjo tool to explore duplications, 
we have asked them to answer some questions related to the experience of using the tool and what could be
improved. The question asked can be found below and the answers can be found in Appendix Y.

>>>>

The ethnographic study happened in two phases. 
In the first phase, the author of this research explored the process of mitigating
a code duplication found in the AMD Display Driver and sent it as a PATCH to iteract with the
Linux community and get their feedback. On a second phase, we proposed to students on the 
course Free Software Development at University São Paulo to mitigate the code duplications found 
by the tool in the IIO subsystems and the AMD Display Driver and pass the process of send the mitigations as PATCH.
(DOUBLE CHECK IF THE STUDENTS HAVE SENT THE PATCHES).

@@

Given the vast number of lines and components in the Linux kernel, analyzing every duplication is impractical. We focused our scope on the AMD Display driver, which, as shown in Figure \ref{fig:relatory_ex}, contains over 20,000 lines of duplicated code at a similarity metric of 100\%.

@@

Ethnography is a recognized qualitative empirical method for understanding people, 
their cultures, and their work practices \citep{bookethno}. It allows researchers 
to gain insights into values, beliefs, and ideas from the perspective of community 
members \citep{ethnosoft}. In our research, refactoring code duplications without 
community interaction would prevent us from understanding whether Linux maintainers 
would accept our approach or identify their concerns. Thus, we will use ethnography, 
positioning ourselves and students as Linux kernel contributors focused on removing 
identified  code duplications.

This section outlines our approach for mitigating code duplication in the Linux kernel
and describes the ethnographic methodology for engagin with the Linux community. Firstly, 
we present the participant observation research method explored by the author. Then, 
we present the methodolody of the experiment done with the students.

@@

Given the tool evaluations that demonstrate the tool capabilites of finding duplications, 
we proceeded to undertand if the duplications found are possible to be mitigate and
the kernel subsystem's maintainers opinions about the duplications and their respective mitigations.
To accomplish this objective, we initialy proposed a participant-observation experiment, 
where we, people that have never contribute to the Linux kernel, 
act as contributors to the AMD Display driver and collect artifacts of our experience in the process.

As a first step of this experiment, we executed the tool in the AMD Display driver codebase and 
manually analysed the biggest duplications in terms of number of lines and raised to 
two duplicated function pairs that we judged interesting to try to mitigate. 
We have choose to approach the biggest duplications as a proof of concept that 
we can make significant impacts. The two duplicated function pairs are shown in Table X.

Given the two duplicated function pairs, we observed that the files that the functions exist contains 
multiple other duplications. Thus, we proposed a simple systematic approach to mitigate all the 
functions duplicated in the context and not only the duplicated function pairs. 
The systematic approach is defined below.

After we approached the duplications with the systematic approach. For the duplicated function
pairs that had successfully being mitigate the duplication, we sent patches to the AMD Display driver
maillist to validate the mitigations and see the maintainers opinion, while documentaion the process 
and our impressions.

\subsubsection{Systematic Approach to Mitigate Code Duplication}

In C programming, communication between source files is achieved by creating header files that specify 
libraries \citep{Cbook}. Since the Linux kernel is primarily written in C, we eliminated code duplications 
by consolidating duplicated code into a single library, which would replace instances of 
duplication across the codebase.

For each duplicated function pair, for each function in the pair, we used the \textit{function information} functionality 
(described in Section \ref{subsec:functioncommand}) to locate all duplications of the function in the codebase, 
as it may have been duplicated more times then the two occurences of the duplicated function pair. 
This strategy resulted in a collection of duplicated functions and corresponding code files.

To identify shared code more effectively, we extended this approach to search for other common 
functions across all collection files. We then applied specific refactoring methods to each shared 
function. If the functions were identical across files, we removed the duplicates and created a single 
function in the library. If modifications existed, we applied case-specific refactoring.

@@

\todo[inline]{Research with students does not look good.}

Posterior of the initial participant observation that we made, we decided to expand the experiment 
by proposing to students mitigate duplication found by the tool and send patches to the kernel community.

@@

For the first alternative, one of the teaching assistants used the Arkanjo tool in the IIO subsystem and
listed interesting duplicated functions pair that the groups can simple mitigate this specific duplication pair, 
and it is up to the students on how
to refactor the code to mitigate the duplications and pass through the process to send patches to the IIO 
maintainers to review. For the second alternative, we propose a similar experience then the 
participant observation method, where the students pass through the whole process of executing the Arkanjo tool
in the AMD display driver, analyze the duplications found, create a patch to mitigate a duplication and sent
the patch to the AMD display driver's maintainers.

@@

For the teaching assistant and the student that executed the Arkanjo tool to explore duplications, 
we have asked them to answer some questions related to the experience of using the tool and what could be
improved. The question asked can be found below and the answers can be found in Appendix Y.
